A core idea for dynamic community detection, among most approaches, is to use the community membership of each vertex from the previous snapshot of the graph, instead of initializing each vertex into singleton communities. One straightforward strategy for dynamic community detection involves leveraging the community memberships of vertices from the previous snapshot of the graph \cite{com-aynaud10, com-chong13, com-shang14, com-zhuang19}, which has been termed as \textbf{Naive-dynamic (ND)}. Alternatively, more sophisticated techniques have been devised to reduce computational overhead by identifying a smaller subset of the graph affected by changes. These techniques include updating only changed vertices \cite{com-aktunc15, com-yin16}, processing vertices in the proximity of updated edges (within a specified threshold distance) \cite{com-held16}, disbanding affected communities into lower-level networks \cite{com-cordeiro16}, or employing a dynamic modularity metric to recompute community memberships from scratch \cite{com-meng16}.

Zarayeneh et al. \cite{com-zarayeneh21} propose the \textbf{Delta-screening (DS)} approach for updating communities in a dynamic graph. This technique examines edge deletions and insertions to the original graph, and identifies a subset of vertices that are likely to be impacted by the change using the modularity objective. Subsequently, only the identified subsets are processed for community state updates, using the Louvain and Smart Local-Moving (SLM) algorithms \cite{com-waltman13}. Recently, Sahu et al. \cite{sahu2024shared} introduced the parallel \textbf{Dynamic Frontier (DF)} approach for the Louvain algorithm. Given a batch update consisting of edge deletions or edge insertions, the DF approach incrementally identifies an approximate set of affected vertices in the graph with a low run time overhead. They also discuss a parallel algorithm for the Delta-screening (DS) approach.

However, to the best of our knowledge, none of the proposed approaches have been extended to the Leiden algorithm.
